---
title: "Stats218_HW3"
author: "Peter Racioppo"
date: "11/28/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 library(sna)
library(network)
library(networkdata)
```

1) Modeling the French Financial Elite: Here we consider a network collected by Charles Kadushin and described in the Kadushin (1990).
He collected data from 127 members of the French financial elite. He used various criteria to determine the top 28 and recorded their who-to-whom responses to questions about who was influential, who were members of the elite and who were friends. He also recorded a large amount of information on their individual backgrounds and characteristics.
We will focus on the (undirected) friendship network. There are many vertex covariates, including:
• prestige: (coded as 0 if respondent has neither a particule nor a social register listing; 1 if a respondent has either a particule or a social register listing; and 2 if respondent has both a particule and social register listing)
• party: An indicator of the party membership. There are 11 parties.
• masons: A member of the masons? 1=no; 2=yes.
• ena: Graduated from ENA? 1=no; 2=yes.
• boards: Number of top boards they are a member of.
Many more are described in the paper. The data are in the networkdata package

```{r eval=TRUE}
library(networkdata)
library(ergm)
data(ffef)
help(ffef)

```

##########################################################

a) Plot the network with the vertex color being the ENA attendance. What do you see?

```{r eval=TRUE}

# Record ena values in array ena
v = ffef$val
ena = 1:length(v)*0
for(i in 1:length(v)){
        x=v[[i]]
        if(x$ena == 1){
                ena[i] = "blue"
        }
        else if(x$ena == 2){
                ena[i] = "red"
        }
        else {
                ena[i] = "black"}
}

plot.network(ffef, vertex.cex = 3,vertex.col = ena)

```

In the above plot, blue represents not graduating from ENA and red represents graduating. Clustering based on ENA value can be easily observed in the plot.

##########################################################

b) Fit a model to the network that includes terms for the homophily on ENA attendance, prestige and party affiliation. Include terms for geometrically weighted edgewise shared partners with the scale parameter fixed at 0.5 (i.e., gwesp(0.5,fixed=T)). Include a similar term for geometrically weighted dyadwise shared partners with the scale parameter fixed at 0.5 (i.e., gwdsp(0.5,fixed=T)).

```{r eval=TRUE}

fit <- ergm(ffef ~ edges + nodematch("ena") + nodematch("prestige") + nodematch("party") + gwesp(0.5,fixed=T) + gwdsp(0.5,fixed=T))
summary(fit)

```

##########################################################

c) Give an interpretation for each of the coefficients in the model in terms of what it means and also what its magnitude indicates about the nature of social relations in the network.

The "edges" term (which indicates the total number of edges in the graph) has a large magnitude, indicating that it is a strong predictor of graph characteristics. It is negative, which indicates that graphs with fewer edges are more likely. The coefficients for "ena," "prestige," and "party," all are positive, indicating that there is homophily (that is, for all three characteristics. The effect is strongest for "ena," then "party" and lastly "prestige." 

Given two nodes A and B, connected by an edge, the edge-wise shared partners are the nodes which are connected by a single edge to either A or B. The dyad-wise shared partners are the same thing, except A and B do not have to be connected by an edge. The geometric ewsp and dwsp measures weight larger neighborhoods geometrically less. These terms are measures of clustering or transitivity.

The MLE fit gives positive coefficients for both measures, but a higher value for the edge-wise fixed partners than the dyad-wise fixed partners.

The edge term, the ena homophily, and the gwdsp terms are the most statistically significant, but all terms are least 0.05 statistically significant.

##########################################################

d) Look at the MCMC diagnostics for the model (via, e.g., mcmc.diagnostics(fit)). What does it say about the convergence of your model?

```{r eval=TRUE}
mcmc.diagnostics(fit)
```

The MCMC error percentages are 0 for all six covariates. The histograms of all six covariates appear to be approaching relatively smooth, Gaussian-like curves, which suggests that the model is converging and has well-structured statistics.

##########################################################

e) Extend the model to include other covariates in the network and other terms that you think are interesting in explaining the social structure. Feel free to consult the reference paper for ideas. Overall, what are the important features of the social structure of this network?

I began by including most of the other covariates. Only ena, party, prestige, birthdate, and normal-sch have statistically significant effects and only ena was highly significant.

```{r eval=TRUE}

fit <- ergm(ffef ~ edges + nodematch("birthdate") + nodematch("birthplace") + nodematch("sciencepoly") + nodematch("polytech") + nodematch("university") + nodematch("normal-sch") + nodematch("ena") + nodematch("inspec-gen") + nodematch("cabinet") + nodematch("finance-min") + nodematch("party") + nodematch("religion") + nodematch("masons") + nodematch("zipcode") + nodematch("socialreg") + nodematch("elitevote") + nodematch("eliteprom") + nodematch("prestige") + nodematch("clubs") + nodematch("topboards"))
summary(fit)

```

I kept only birthdate, ena, party, and prestige. I tried fitting a model for differential homophily on the covariates, but this did't appear to have a big effect. I also tried nodefactor, nodecov, and absdiff, which were also of no use. I also tried nodemix("ena"), which gave no significant improvement.


```{r eval=TRUE} 

fit <- ergm(ffef ~ edges + nodematch("birthdate") + nodematch("ena") + nodematch("party") + nodematch("prestige") + gwesp(0.5,fixed=T) + gwdsp(0.5,fixed=T))
summary(fit)

```

It appears that the important features of this social network are degree, ena, birthdate, party, prestige, and transitivity (as measured by gwesp and gwdsp).

##########################################################

2) Balance in signed networks: Consider the alliance structure among three tribal groups of the Eastern Central Highlands of New Guinea. considered by Read (1954). Among these groups (the Gahuku-Gama) the enemy of an enemy can be either a friend or an enemy.
The data is in the networkdata package as a combination of the rona and hina networks. The rova binary network has 1 for alliance (“rova”) relations and 0 otherwize. The hina binary network has 1 for antagonistic (“hina”) relations. The default (non)edge value of 0 means neither of these relations hold.

##########################################################

a) First we construct a valued network to measure the degree of alliance on a count scale. This will have 2 for alliance (“rova”) relations, 0 for antagonistic (“hina”) relations, and 1 for neither of these. Store these in the edge attribute count. This representation reduces the signed network onto an ordinal alliance scale.
Here is some code to do it:

```{r eval=TRUE}
# library(ergd)
library(ergm.count)

data(rova)
data(hina)
gama <- rova[,]+(1-hina[,])
diag(gama) <- 0
gama <- as.network(gama, directed = FALSE, matrix.type = "a",
ignore.eval = FALSE, names.eval = "count")
```

In this analysis we will treat the value as a binomial count with two trials and fit ERGM via the ergm.count package.

##########################################################

b) Plot the network, labeling the nodes and using the edge.col parameter. Briefly describe the structure you see.

```{r eval=TRUE}

# Record ena values in array s
v = gama$val
names = 1:length(v)*0
for(i in 1:length(v)){
        x=v[[i]]
        names[i] = x[2]
}

# Record the edge values in array s
s = rova[,]+(1-hina[,])
diag(s) <- 0
s2 = s+2

plot.network(gama, vertex.cex = 3,vertex.col = "black",edge.col = s2,label=names)

gama_dir <- as.network(s, directed = TRUE, matrix.type = "a", ignore.eval = FALSE, names.eval = "count")

```

There appears to be clustering based on edge value. Tribes with the most number of alliances seem to be more central, while tribes with many antagonisms are more outlying.

##########################################################

c) Fit a model with the nonzero and sum terms. Interpret what the terms in the model mean and what their estimates mean for the relations network. Hint: See help("ergm-terms").

```{r eval=TRUE}

# fit <- ergm(gama_dir ~ edges + absdiff(s,"sum"))
# fit <- ergm(gama_dir ~ edges + edgecov(s, "nonzero") + edgecov(s, "sum"))
# fit <- ergm.count(gama ~ edges + CMP)

# fit <- ergm.count(gama ~ edges + dyadcov(s, "nonzero") + dyadcov(s, "sum"))
fit <- ergm(gama ~ edges + edgecov(s, "nonzero") + edgecov(s, "sum"))
summary(fit)

```

The nonzero edge covariate is higher-valued than the edges statistic.

##########################################################

d) Add to this model the term transitiveties(threshold=1.5). Give an interpretation of this term and what the estimates of each of the terms mean for the relations network.

```{r eval=TRUE}

# fit <- ergm.count(gama_dir ~ edges + transitiveties(threshold=1.5))
# summary(fit)
# This returns the following error:
# <<Error in ergm.count(gama_dir ~ edges + transitiveties(threshold = 1.5)) : 
# could not find function "ergm.count">>

# fit <- ergm(gama_dir ~ edges + transitiveties(threshold=1.5))
# summary(fit)
# This returns the following error:
# <<Error: In term ‘transitiveties’ in package ‘ergm’: Model term does not recognize ‘threshold’ argument.>>

# Instead, I attempted to use the related transitive, ttriple, and ctriple statistics:
# fit <- ergm(gama_dir ~ edges + transitive)
fit <- ergm(gama_dir ~ edges + ttriple)
# fit <- ergm(gama_dir ~ edges + ttriple + ctriple)
summary(fit)

```
The "transitiveties" term  adds a statistic on the number of paths of length 2 in the network. On the other hand, the "ttriples" term adds a statistic on the number of transitive triples in the network, and "ctriples" counts the cyclic tripels. The "transitive" term adds a statistic on the number of triads in the network that are transitive. None of the fits using these terms was statistically significant.

##########################################################

e) See if you can extend the model by adding further terms. Briefly report if you can find a model that fits better than the models in parts a) and b).

```{r eval=TRUE}
fit <- ergm(gama_dir ~ edges + ttriple + gwesp)
summary(fit)

```             
Adding a gwesp term greatly improved convergence. The model now converges, and the ttriple term is now statistically significant.

##########################################################

f) Finally, we have represented the signed network on an ordinal alliance scale. Comment on what is lost by this, if anything, compared to the signed network.

No information is lost by converting the signed network to an ordinal network, because each of the possible relations in the signed network can be represented in the ordinal network. However, since these relationships are represented on a count scale, "no alliance" is represented as taking an intermediate value (1) between "alliance" (2) and "antagonism" (0), which is different than in the signed graph.

##########################################################

3) Modeling a Triad Census in Friendship Relations: Here we consider again the network introduced in Homework 2 of strong friendship ties among 13 boys and 14 girls in a sixth- grade classroom, as collected by Hansell (1984). Each student was asked if they liked each other student “a lot”, “some”, or “not much”. Here we consider a strong friendship tie to exist if a student likes another student “a lot.” Also recorded is the sex of each student. The data is in the networkdata package

```{r eval=TRUE}
detach("package:ergm.count", unload=TRUE)
library(networkdata)
library(ergm.tapered)
data(hansell)
help(hansell)
```

##########################################################

a) Fit a triad census model using ergm.tapered. Include a term for homophily by sex.

```{r eval=TRUE}

fit <- ergm.tapered(hansell ~ edges + nodematch("sex") + triadcensus(1:7))
summary(fit)

```
##########################################################

b) Give a brief interpretation of the coefficients of the terms. Based on this model, does there appear to be a general preference for transitive friendship ties? Is there homophily by sex? Give a brief summary of the overall triad census pattern.

The homophily coefficient is positive, has the largest magnitude of any coefficient, and is highly statistically significant, so we can conclude that there is homophily by sex. The edges statistic is negative and also highly statistically significant. The 021C triad coefficient is highly statistically significant and negative. This suggests that there is a significant preference against 021C triads, which can be thought of as hierarchies of length 2, so the generative model for the network likely has an hierarchical structure. In fact, 5 of the 7 triad statistics are statisistically significant, and they are all negative. In summary, the model suggests there is a preference against transitive friendship terms.

##########################################################

4) Modeling a Triad Census in Friendship Relations: Here we consider the network of friendship ties of girls in grade 9. The data is in the networkdata package.

```{r eval=TRUE}

# detach("package:ergm", unload=TRUE)
library(networkdata)
data(gfriends)
help(gfriends)

```

##########################################################

a) First, extract out the grade nine students using gfriends$X[,1]==9 and create a directed network. Also add a vertex variable equal to the student gpa.

```{r eval=TRUE}

grade9 = (gfriends$X[,1]==9)*1
idx = which(grade9==1)
gpa = gfriends$X[,2]
gpa9 = gpa[idx]
smokes = gfriends$X[,3]
smokes9 = smokes[idx]
adj = gfriends$Y
adj9 = adj[idx,idx]

gf <- network(x = adj9,directed = TRUE,loops = FALSE,matrix.type = "adjacency")
set.vertex.attribute(gf, # the name of the network object
                     "gpa", # the name we want to reference the variable by
                     gpa9 # the value we are giving that variable
                     ) 
set.vertex.attribute(gf, # the name of the network object
                     "smoke", # the name we want to reference the variable by
                     smokes9 # the value we are giving that variable
                     )

gplot(gf)

```

Fit a triad census model using a subset of the first seven triad types (012, 102, 021D, 021U, 021C, 111D, 111U). Recall that 003 is the reference category. Also include GPA and smoking effects, using something like:

```{r eval=TRUE}

# library(ergd)
library(ergm.tapered)
fit <- ergm.tapered(gf ~ edges + nodecov("gpa") + nodecov("smoke") + triadcensus(c(1:7)))
summary(fit)

```

##########################################################

b) Give a brief interpretation of the coefficients of the terms. Give a brief summary of the overall triad census pattern.

There is a small (though not statistically significant) dependence on the node covariates, with a positive relationship for gpa and negative for smoking. Most of the triad coefficients are negative and not statistically significant.

The two triad types which are most significant and have the largest magnitude are 021D and 111U. The 021D triad is a directed network with two edges pointing from a single node into the other two nodes in the triad. A preponderance of triads of this type would indicate that the data is hierarchical. The 111U triad has one multi-directional edge and a directed edge leaving one of the nodes connected to the multi-directional edge. In other words, the 111U triad is like the 021D triad, but with an additional directed edge pointing back along one of the edges. Thus, a network with many 111U triads can be thought of as a hierarchical directed network, but with some edges pointing back up the hierarchy. That is, twice as many edges point down the hierarchy as point up.

In summary, since the coefficients of both the 021D triad and the 111U are negative, high in magnitude, and statistically significant, we can conclude that the generative model for this network has a preference for non-hierarchical relationships.

##########################################################

c) Look at the MCMC diagnostics for the model (via, e.g., mcmc.diagnostics(fit)). What does it say about the convergence of your model?

```{r eval=TRUE}
mcmc.diagnostics(fit)
```

All of the statistics in this model follow distributions that appear to be roughly Gaussian. They are all approximately zero-mean and symmetric, although the distributions for triads 102, 021D, and 021U appear to be slightly skewed left with slightly long tails. Sparse, locally dependent random graph models have statistics that are asymptotically Gaussian. In view of the closeness of the model statistics to Gaussian distributions, we can conclude that the model is converging.

##########################################################

d) Try to fit a model with fewer terms than that in a). What is the preferred model?

The original model had an AIC of 441.2 and a BIC of 488.2, with statistically significant 012D and 111U triad terms.

I began by removing the two least statistically significant triad terms: triads 012 and 102. This fit gave an AIC of 422.4 and a BIC of 460, with statistically significant 021D and 021C triad terms and edges.

I then removed triad 021U. This gave an AIC of 422.2 and a BIC of 455.1, with statisically significant 021D, 021C, 111U triad terms, edges, and gpa covariate.

I then removed triad 111D, which gave an AIC of 418 and a BIC of 446.2. Removing the smoking covariate gave a model with an AIC of 414.5 and BIC of 438. Finally, removing the 111U triad gave an AIC of 415.8 and BIC of 434.6.

The final model includes edges, the GPA covariate, and triads 021D and 021C. All terms are statistically significant. There is a positive dependence on edges and GPA and negative dependence on the triad terms. The MCMC diagnostics look good (Gaussian).


```{r eval=TRUE}

library(ergm.tapered)
fit <- ergm.tapered(gf ~ edges + nodecov("gpa") + triadcensus(c(3,5)))
summary(fit)

```
```{r eval=TRUE}
# mcmc.diagnostics(fit)
```

##########################################################

5) Model for Protein-protein interaction data: Butland et al (2005) “Interaction network containing conserved and essential protein complexes in Escherichia coli” reported a net- work of protein-protein interactions (bindings) that we obtained from http://pil.phys. uniroma1.it/~gcalda/cosinsite/extra/data/proteins/. This data is available in the networkdata package

```{r eval=TRUE}
library(networkdata)
data(butland_ppi)
help(butland_ppi)
```

##########################################################

Convert the edgelist to a directed network (The el2sm function may be helpful).

```{r eval=TRUE}

adj = el2sm(butland_ppi)
bl <- network(x = adj,directed = TRUE,loops = FALSE,matrix.type = "adjacency")
gplot(bl)

```

##########################################################

Fit various tapered ERGM models to the network using ergm.tapered. Consider terms documented under ergm-terms. Good candidates include istar, ostar, gwodegree, gwidegree, dgwest, dgwdsp, ctriple, ttriple.

I began by fitting a model on istars and ostars from 1 to 3. The models did not converge.

I next fit models on gwodegree and gwidegree, the geometrically-weighted degree distributions. These also did not converge.

I next fit a dgwdsp term (dgwdsp is the geometrically-weighted edgewise shared partner distribution). This also did not converge.

I next fit a model on ttriple, which did converge. The edges and ttriple statistics are both highly statistically significant. The edges coefficient is negative and larger in magnitude, while the ttriple statistic is positive, which indicates a tendency to form transitive triads.
On the other hand, ctriple did not converge.

I next fit a model on ttriple and istar(1), which did not converge.

```{r eval=TRUE}

library(ergm.tapered)
# fit <- ergm.tapered(bl ~ edges + istar(1))
# fit <- ergm.tapered(bl ~ edges + gwodegree(1))
# fit <- ergm.tapered(bl ~ edges + dgwdsp)
fit <- ergm.tapered(bl ~ edges + ttriple)
summary(fit)

```
##########################################################

Check the MCMC diagnostics with mcmc.diagnostics. Overall, how does the goodness-of-fit look?

```{r eval=TRUE}
mcmc.diagnostics(fit)
```
Th edges and ttriple statistics are not perfectly symmetric, but nevertheless they are roughly Guassian, which indicates the model is approaching convergence.
